---
title: "Assignment II: Group C"
author: "Lam Ching Yin"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Name**]{style="color:red"}**: Lam Ching Yin**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.*
:::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

-   `ggplot2::msleep`
:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

:::: callout
### Your answer is here

::: callout
The `msleep` (Mammalian Sleep) dataset contains the sleep habits and physiological characteristics of **83 different species of mammals**. It is primarily used to explore the relationship between an animal's body weight, brain weight, and diet relative to how much sleep they require.

The data set contains the following attribute: name, genus, vore, order, conservation, sleep_total, sleep_rem, sleep_cycle. awake. brainwt(brain weight), bodywt(body weight).

The data in `msleep` is a **compiled dataset**, meaning it wasn't collected in a single experiment. Instead, it was aggregated from various primary research sources.

**Sleep Data:** These values were typically measured using **Electroencephalography (EEG)** in laboratory settings to distinguish between REM and non-REM sleep cycles.

**Weights:** Brain and body weights were recorded from anatomical studies and biological surveys.

**Compilation:** Savage and West updated the 1976 Allison and Cicchetti data by cross-referencing newer literature to fill in gaps and add more species

Additional variables order, conservation status and vore were added from wikipedia.
:::
::::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# YOUR CODE HERE
# xxxxx
```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
### Your answer is here

xxxxxxxxxxxxxx
:::

```{r}
library(ggplot2)

# View the first 6 rows
head(msleep)

# Get a summary of the statistics
summary(msleep)

# Check the data types of each column
str(msleep)
```
