---
title: "Assignment II: Group C"
author: "Chan Chi Him"
format: 
  html:
    toc: true
    code-fold: false
    embed-resources: true # <= check here: https://quarto.org/docs/reference/formats/html.html
execute:
  echo: true    # <= knitr::opts_chunk$set(echo = TRUE) in Rmarkdown
  warning: false
  message: false
---

# Name

[**Chan Chi Him**]{style="color:red"}**: 21064932**

# Generative AI declaration

Students may use generative AI tools during assignment, for example, to check and improve code, provided that the content is original and reflects the student's own contributions. If generative AI is used, students must declare its use in the box below. An example statement is: “*During the assignment, I used \[XXXX\] to \[XXXX\]. I subsequently reviewed and edited the code and take(s) full responsibility for the final version of the code.”*

::: callout
In this assignment, I use AI to modify my codes, checking whether there are any mistakes and ask it to help me fix the problem. :::

# Dataset

Each group member may choose either of the following dataset:

**Built-in dataset**

-   `palmerpenguins::penguins`
-   `ggplot2::msleep`
-   `ggplot2::midwest`
-   `datasets::quakes`

**Public dataset**

BIOTIME (CC0) <https://biotime.st-andrews.ac.uk/usageGuidelines.php>

-   `BIOTIME_raw_data_192.csv` (uploaded on Canvas) (Whale data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=192>)
-   `BIOTIME_raw_data_232.csv` (uploaded on Canvas) (Fish data from <https://biotime.st-andrews.ac.uk/selectStudy.php?study=232>)

# Figures by each group member

### Dataset

::: callout-important
[Which dataset did you choose?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
`datasets::quakes`


:::

### Introduction

::: callout-important
[What is the dataset about? What variables are included in the data? How was the data taken/measured?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
The dataset is about the data recorded from 200 earthquake events. In the dataset, there are five variables, which is the location of latitude and longtitude that the earthquake happens, depth that measures how deep does the earthquake goes, magnitude that means the strength of the earthquake and stations that means the number of different seismic stations that detected and reported the earthquake. 
:::

### Interactive plot(s) using `plotly`

::: callout-important
[Please explain which plot type you choose and why. Also, please discuss the strengths and weaknesses of your plot(s).]{style="color:darkred; font-weight:bold;"}
:::

::: callout
In the Fiji region (where this data originates), the Pacific Plate is subducting under the Australian Plate. A 2D map would only show a cluster of dots; a 3D plot allows you to see the angle of the subduction zone as the earthquakes get deeper the further they move from the fault line.
For the strengths:
1. Visualizes the Subduction Slab: By reversing the Z-axis, it mimics the real-world Earth interior.
2. Multivariate Depth: Using a color gradient (e.g., Viridis) makes it immediately obvious which quakes are shallow (crust) versus deep (mantle).

For the weakness:
1. Overcrowding: With 1,000 observations in the full quakes dataset, the markers can overlap, making it hard to see individual points in the middle of the cluster.
2. Depth Perception: Without rotating the plot, it can be difficult to tell exactly where a point lies in 3D space.

:::

Note: At least one plot is required, but multiple plots are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
The primary purpose of this 3D visualization is to reveal the spatial architecture of earthquake activity, specifically identifying the presence of a subduction zone. While a flat map tells you where people felt the ground shake, this 3D plot tells you the geological story of what is happening miles beneath the surface.

xxxxxxxxxxxxxx
:::

Note: If you want to specify setting for a code chunk, you can use `#|` for each option (for example, `#| echo: true`). Rmarkdown code chunk still works (for example, `{r echo = true}`), but Quarto style is recommended.

```{r}
#| echo: true
# 1. Load the libraries
library(ggplot2)
library(gganimate)
library(viridis)

# 2. Create the plot using the built-in quakes dataset
df <- datasets::quakes

# We use shadow_mark() to "build" the map from shallow to deep
p <- ggplot(df, aes(x = long, y = lat, color = depth, size = mag)) +
  geom_point(alpha = 0.6) +
  # Using 'option = "D"' provides the classic Viridis scale
  scale_color_viridis_c(option = "D", direction = -1) +
  theme_minimal() +
  labs(title = 'Fiji Earthquakes at Depth: {frame_time} km',
       subtitle = 'Visualizing the descending tectonic plate',
       x = 'Longitude', y = 'Latitude',
       color = 'Depth (km)', size = 'Magnitude')

# 3. Define the animation transition
anim <- p + 
  transition_time(depth) + 
  shadow_mark(past = TRUE) # Keeps older points visible as the "sink" continues

# 4. Render the result
animate(anim, nframes = 100, fps = 10, renderer = gifski_renderer())
```

### Animation using `gganimate`

::: callout-important
[Please explain which animation type you choose and why. Also, please discuss the strengths and weaknesses of your animation.]{style="color:darkred; font-weight:bold;"}
:::

::: callout
I have chosen a Cumulative Point Animation that reveals earthquakes from the surface (shallow) down to the deepest point (680 km). Since instead of showing all 1,000 points at once, animating by depth allows the viewer to see the "shape" of the tectonic plate as it descends. By using shadow_mark(), we leave the previous points on the screen, gradually building a map of the subduction zone. It mimics the geological process of the plate "diving" into the mantle.
For the strength:
1. Reduced Cognitive Load: By showing the quakes in stages, the viewer isn't overwhelmed by 1,000 dots simultaneously.
2. Narrative Flow: It tells a story of depth. You can clearly see that shallow quakes cover a wider area, while deep quakes are concentrated in a specific linear "trench" or slab.

For the weakness:
1. Loss of 3D Perspective: Unlike the plotly version, gganimate produces a 2D file (GIF/MP4). We lose the ability to rotate the plot.
2. File Size: High-quality GIFs with 1,000 points and many frames can be large and slow to render.
:::

Note: At least one animation is required, but multiple animations are allowed

::: callout-important
[What is the purpose of this visualization?]{style="color:darkred; font-weight:bold;"}
:::

::: callout
The purpose of this visualization is to reveal the internal geometry of the Earth, specifically illustrating the subduction process occurring in the Fiji region
While a standard list of numbers tells you how many earthquakes happened, an animated visualization converts those numbers into a geological "story."
:::

```{r}
# 1. Load the libraries
library(ggplot2)
library(gganimate)
library(viridis)

# 2. Create the plot using the built-in quakes dataset
df <- datasets::quakes

# We use shadow_mark() to "build" the map from shallow to deep
p <- ggplot(df, aes(x = long, y = lat, color = depth, size = mag)) +
  geom_point(alpha = 0.6) +
  # Using 'option = "D"' provides the classic Viridis scale
  scale_color_viridis_c(option = "D", direction = -1) +
  theme_minimal() +
  labs(title = 'Fiji Earthquakes at Depth: {frame_time} km',
       subtitle = 'Visualizing the descending tectonic plate',
       x = 'Longitude', y = 'Latitude',
       color = 'Depth (km)', size = 'Magnitude')

# 3. Define the animation transition
anim <- p + 
  transition_time(depth) + 
  shadow_mark(past = TRUE) # Keeps older points visible as the "sink" continues

# 4. Render the result
animate(anim, nframes = 100, fps = 10, renderer = gifski_renderer())
```
